---
title: "skiPatrol Quickstart"
subtitle: "Connect, query, write tables, and bridge to DBI/dbplyr"
format: html
---

A first tour of `skiPatrol`: connect to Snowflake, run queries, move a
`data.frame` in and out of a table, and hand off to `skiLift` when you want
DBI or `dbplyr`.

**Runs unmodified in three environments** — a local R session (RStudio,
Positron, VS Code, terminal), the Posit Native App, and Snowflake Workspace
Notebooks — via `sfr_connect()`'s auto-detection. Configure it, if you need
to at all, through `_config.R`.

**What you'll do:**

1. Connect
2. Run queries and move data
3. Aggregate in Snowflake rather than in R
4. Bridge to DBI and `dbplyr` via `skiLift` *(optional)*
5. Plot the results
6. Clean up

## Installation

One time only, and not needed at all in a Workspace Notebook or the Posit
Native App if `skiPatrol` is already installed there.

```r
# install.packages("pak")
pak::pak("posit-dev/skiPatrol")

# skiPatrol talks to the Snowflake ML platform through Python. This installs
# the Python side; it is a no-op if the dependencies are already present.
skiPatrol::sfr_install_python_deps()
```

## 1. Connect

```{r}
#| label: setup
library(skiPatrol)
source("_config.R")
```

`sfr_connect()` works out how to authenticate on its own — a local
`connections.toml` profile, the Native App's OAuth profile, or the ambient
Workspace Notebook session. You do not need to tell it which environment it
is in.

Database, schema and warehouse *are* passed explicitly, in every environment.
That is deliberate rather than defensive: several skiPatrol calls (notably
`sfr_write_table()`) issue unqualified SQL that depends on the session's
*current* database, and auto-detection alone does not always set one.

```{r}
#| label: connect
conn <- sfr_connect(
  name      = SFR_CONNECTION_NAME,
  warehouse = SFR_WAREHOUSE,
  database  = SFR_DATABASE,
  schema    = SFR_SCHEMA
)
conn
```

## 2. Queries and table operations

`sfr_query()` returns a plain `data.frame`.

```{r}
#| label: query
sfr_query(conn, "SELECT CURRENT_TIMESTAMP() AS now, CURRENT_USER() AS user_name")
```

Table names go through `sfr_fqn()`, which qualifies a bare name against the
connection's current database and schema. Fully qualified names are used
throughout these examples so that a document behaves the same way whether or
not the session happens to have a database set.

```{r}
#| label: write-table
sfr_write_table(conn, sfr_fqn(conn, "SFR_MTCARS"), mtcars, overwrite = TRUE)
```

```{r}
#| label: list-tables
head(sfr_list_tables(conn), 10)
```

```{r}
#| label: read-table
df <- sfr_read_table(conn, sfr_fqn(conn, "SFR_MTCARS"))
head(df, 5)
```

```{r}
#| label: describe
sfr_list_fields(conn, sfr_fqn(conn, "SFR_MTCARS"))
```

```{r}
#| label: exists
sfr_table_exists(conn, sfr_fqn(conn, "SFR_MTCARS"))
```

## 3. Aggregate in Snowflake

Push the work to the warehouse and bring back only the result. Note that
column names come back uppercase — that is Snowflake's convention, not
something skiPatrol imposes.

```{r}
#| label: aggregate
sfr_query(conn, paste(
  "SELECT CYL, COUNT(*) AS n, AVG(MPG) AS avg_mpg, AVG(HP) AS avg_hp",
  "FROM", sfr_fqn(conn, "SFR_MTCARS"),
  "GROUP BY CYL ORDER BY CYL"
))
```

## 4. DBI and dbplyr via skiLift *(optional)*

`skiPatrol` covers the ML platform; `skiLift` is the companion package that
provides DBI-compliant database access, `dbplyr` translation, and the
RStudio/Positron Connections Pane. `sfr_dbi_connection()` hands you a
`skiLift` connection built from the `sfr_connection` you already have, so you
do not authenticate twice.

You can also create one directly with
`DBI::dbConnect(skiLift::Snowflake(), ...)` if you have no `skiPatrol`
connection to begin with.

```{r}
#| label: dbi
#| eval: !expr requireNamespace("skiLift", quietly = TRUE)
library(skiLift)
library(DBI)

dbi_con <- sfr_dbi_connection(conn)

dbGetQuery(dbi_con, "SELECT 42 AS answer")
```

```{r}
#| label: dbplyr
#| eval: !expr requireNamespace("skiLift", quietly = TRUE) && requireNamespace("dbplyr", quietly = TRUE)
library(dplyr)
library(dbplyr)

# A lazy reference -- no data moves until collect().
cars_tbl <- tbl(dbi_con, I(sfr_fqn(conn, "SFR_MTCARS")))

cars_tbl |>
  group_by(CYL) |>
  summarise(
    n       = n(),
    avg_mpg = mean(MPG, na.rm = TRUE),
    avg_hp  = mean(HP, na.rm = TRUE)
  ) |>
  arrange(CYL) |>
  collect()
```

## 5. Plot

```{r}
#| label: plot
#| fig-width: 7.29
#| fig-height: 4.69
library(ggplot2)

ggplot(df, aes(x = WT, y = MPG, color = factor(CYL))) +
  geom_point(size = 3) +
  labs(
    title = "MPG vs weight by cylinder count",
    x     = "Weight (1000 lbs)",
    y     = "Miles per gallon",
    color = "Cylinders"
  ) +
  theme_minimal()
```

## 6. Clean up

Left commented so that running the whole document twice does not delete
anything unexpectedly.

```{r}
#| label: cleanup
# sfr_execute(conn, paste("DROP TABLE IF EXISTS", sfr_fqn(conn, "SFR_MTCARS")))
# sfr_disconnect(conn)
```

## Where to go next

| Document | Covers |
|---|---|
| `feature-store.qmd` | Entities, feature views, training data |
| `model-registry.qmd` | Logging, versioning, and deploying R models |
| `model-monitoring.qmd` | Drift and performance monitoring |
| `forecasting.qmd` | An ARIMA worked example, end to end |
